---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 300,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->

<h1 align=center><code>geosx</code></h1>

<p align="center">
    <!-- <a href="https://github.com/atsyplenkov/geosx/releases"><img src="https://img.shields.io/github/v/release/atsyplenkov/geosx?style=flat&labelColor=1C2C2E&color=198ce7&logo=GitHub&logoColor=white"></a> -->
    <!-- <a href="https://cran.r-project.org/package=geosx">
        <img src="https://img.shields.io/cran/v/geosx?style=flat&labelColor=1C2C2E&color=198ce7&logo=R&logoColor=white"></a> -->
    <a href="https://app.codecov.io/gh/atsyplenkov/geosx"><img src="https://img.shields.io/codecov/c/gh/atsyplenkov/geosx?style=flat&labelColor=1C2C2E&color=256bc0&logo=Codecov&logoColor=white"></a>
    <a href="https://github.com/atsyplenkov/geosx/actions/workflows/R-CMD-check.yaml"><img src="https://img.shields.io/github/actions/workflow/status/atsyplenkov/geosx/R-CMD-check.yaml?style=flat&labelColor=1C2C2E&color=256bc0&logo=GitHub%20Actions&logoColor=white"></a>
</p>
<!-- badges: end -->

The `geosx` package provides a collection of common tools for
vector geometry operations as an extension to
[`geos`](https://github.com/paleolimbot/geos). It is a highly
experimental and opinionated library, with a focus on performance, 
simplicity and lightweightness. I originally put this package together for my
own projects, collecting handy functions I kept reusing.

> [!note]
> This package is still in development and API is subject to change.

## Installation

You can install the development version of geosx like so:

```r
# install.packages("pak")
pak::pak("atsyplenkov/geosx")
```
## Examples

For example, one can generate a grid of polygons over some points and
find the points within the central grid cell, or find standalone points
(those which stand within a user-defined distance from other points):
```{r example}
library(geos)
library(geosx)

# Generate some random points
pts <- wk::xy(
  x = runif(10000, 0, 100000),
  y = runif(10000, 0, 100000)
) |>
  geos::as_geos_geometry()

# Generate a grid of polygons over the points
grid <- geos_make_grid(pts, 5, 5)

# Find the points within the central grid cell
pts_within <- geos_str_intersection(pts, grid[13])

# Or find standalone points (those which stand within `maxdist`)
pts_standalone <- geos_standalone_points(pts, 1000)

# Plot the points and the grid
plot(pts, col = "grey")
plot(pts_within, add = TRUE, col = "dodgerblue", pch = 19)
plot(pts_standalone, add = TRUE, col = "forestgreen", pch = 19)
plot(grid[13], add = TRUE, border = "red", lwd = 3)
```

The `geos_str_intersection` function acts similarly to the
`geos::geos_intersect` (and `sf::st_intersection`), but is
slightly faster for multiple polygons. However, one may not find
significant speed gain in single polygon cases due to additional type checking
going under the hood in `geosx`.

```{r benchmark1}
library(bench)
library(sf)

pts_sf <- sf::st_as_sfc(pts)
grid_sf <- sf::st_as_sfc(grid)

# Single polygon
bench::mark(
  geosx = geos_str_intersection(pts, grid[13]),
  geos = pts[geos::geos_intersects(pts, grid[13])],
  sf = sf::st_intersects(pts_sf, grid_sf[13]),
  iterations = 30L,
  check = FALSE
)

# Multiple polygons
bench::mark(
  geosx = geos_str_intersection(pts, grid[13:15]),
  geos = pts[geos::geos_intersects(
    pts,
    geos::geos_make_collection(grid[13:15])
  )],
  sf = sf::st_intersects(pts_sf, grid_sf[13:15]),
  iterations = 30L,
  check = FALSE
)
```

And my favourite tool is the `geos_clip` function, which clips a geometry to a polygon. 

```{r benchmark2}
bench::mark(
  geosx = geos_clip(pts, grid[13]),
  sf = sf::st_intersection(pts_sf, grid_sf[13]),
  iterations = 30L,
  check = FALSE
)
```